---
title: "BiometRIC data analysis"
output: html_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

#####Contact: 

Lu Zhang, lu.zhang@sonova.com for clinical data questions.

## Introduction
This document hosts the code of BiometRic data analysis. Please set working directory of data files before calling functions.

```{r setup}
knitr::opts_knit$set(root.dir = "C:/Users/57lzhang/Desktop/R code for Data Science/Gen2 QoS Summary")
```

###Call all packages
```{r package,echo=TRUE,warning=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
library(MLmetrics)
library(ggplot2)
library(gridExtra)
library(stringr)
library(gmodels)
library(caret)
library(psych)
library(formattable)
```


###Accuracy Table
```{r accuracy table,echo=TRUE,warning=FALSE,message=FALSE}
table_all <- function(){
  temp <- list.files(pattern="^PTek*")
  df <- data.frame()
  for (i in temp){
    ##substract subject id and activity type from filename
    tt<-str_split(i,'-')[[1]][3]
    Subject<-str_split(tt,' ')[[1]][2]
    Subject<-str_c("F-",Subject)
    Activity<-str_split(tt,' ')[[1]][3]
    ##label data with subject and activity
    t<-read.csv(i,header = T)
    t1<-t%>%mutate(Subject = Subject)%>%mutate(Activity = Activity)
    t1 <- subset(t1, select = -c(RRi,RRi.1))
    t1<-t1%>%slice(32:212)
    ##combine all files into one
    df <- bind_rows(t1,df)
  }
  df <- df%>%mutate(Diff = HR.1 - HR)
  table <- df%>%group_by(Activity)%>%drop_na(HR.1,HR)%>%
    summarise(Counts = round(length(HR.1)),
              Mean_Bio = round(mean(HR.1),1),
              SD_Bio = round(sd(HR.1),1),
              Mean_Polar = round(mean(HR),1),
              SD_Polar = round(sd(HR),1),
              Pearson_Corr = round(cor(HR,HR.1),2),
              Bias = round(mean(Diff),1),
              MAE = round(MAE(HR.1,HR),1),
              MAPE = round(MAPE(HR.1,HR)*100,1)
             # Bias_CI_l = round(ci(Diff, 0.95, 0.05,na.rm=T)[2],1),
            #Bias_CI_h = round(ci(Diff, 0.95, 0.05,na.rm=T)[3],1)
    )
  
  ##calculate ICC
  icc_all <-data.frame()
  for (i in c("housework","running","biking","walking","drinking","chewing","talking","sitting")){
    icc<-round(ICC(df%>%filter(Activity == i)
                   %>%select(HR,HR.1)%>%drop_na())$results[3,2],2)
    icc_ci_low <- round(ICC(df%>%filter(Activity == i)
                            %>%select(HR,HR.1)%>%drop_na())$results[3,7],2)
    icc_ci_high <- round(ICC(df%>%filter(Activity == i)
                             %>%select(HR,HR.1)%>%drop_na())$results[3,8],2)
    df_icc<-data.frame(Activity = i,ICC = icc) #ICC_CI_l = icc_ci_low, ICC_CI_h = icc_ci_high 
    icc_all <- bind_rows(df_icc,icc_all)
  }
  
  ##join two tables
  full_table<-left_join(table,icc_all,by="Activity")%>%arrange(desc(ICC))
  ##calculate average of all stats
  summary_row<-summarise_each(full_table,funs(mean))
  
  return(list(
    "table" = formattable(full_table, align = c("l",rep("c", ncol(full_table) - 5),"r","r","r","r"),list(
    `Activity` = formatter("span", style = ~ style(color ="grey",
                                                   font.weight = "bold")),
    `ICC` = color_bar("#FA614B"),
    `MAPE` = color_bar("#B1CBEB"),
    `MAE` = color_bar("#71CA97"),
    `Bias` = formatter("span",style = ~ style(color = ifelse(`Bias`> 0, "green","red")))

    )),
              
    "summary" = formattable(round(summary_row,2),align = c("l",rep("c", ncol(summary_row) - 1)),list(
    `Activity` = formatter("span", style = ~ style(color ="grey",
                                                   font.weight = "bold"))
    ))
    ))
}
```
Call table_all function
```{r,table call,echo=TRUE,warning=FALSE,message=FALSE}
calltable<-table_all()
calltable$table
calltable$summary
```


###Boxplot
boxplot_all function returns two plots - mae vs activties, mape vs activities of all subjects statistics
```{r boxplot,echo=TRUE,warning=FALSE,message=FALSE}
boxplot_all <- function(){
  temp <- list.files(pattern="^PTek*")
  df <- data.frame()
  for (i in temp){
    ##substract subject id and activity type from filename
    tt<-str_split(i,'-')[[1]][3]
    Subject<-str_split(tt,' ')[[1]][2]
    Subject<-str_c("F-",Subject)
    Activity<-str_split(tt,' ')[[1]][3]
    ##label data with subject and activity
    t<-read.csv(i,header = T)
    t1<-t%>%mutate(Subject = Subject)%>%mutate(Activity = Activity)
    t1 <- subset(t1, select = -c(RRi,RRi.1))
    t1<-t1%>%slice(32:212)
    ##combine all files into one
    df <- bind_rows(t1,df)
  }
  df$Activity <- factor(df$Activity,levels = c("sitting", "talking", "chewing", "drinking", "walking", "biking","running","housework"))
  mae_act<-df%>%group_by(Activity,Subject)%>%drop_na(HR.1,HR)%>%summarise(MAE = MAE(HR.1,HR))
  mape_act <-df%>%group_by(Activity,Subject)%>%drop_na(HR.1,HR)%>%summarise(MAPE = MAPE(HR.1,HR)*100)
  mae_boxplot<-ggplot(mae_act,aes(Activity,MAE,fill = Activity))+geom_boxplot()+ylab("MAE (bpm)")+ggtitle("MAE vs Activity in all Subjects")+ylim(c(0,50))
  mape_boxplot<-ggplot(mape_act,aes(Activity,MAPE,fill = Activity))+geom_boxplot()+ylab("MAPE (%)")+ggtitle("MAPE vs Activity in all Subjects")+geom_hline(yintercept = 10,linetype="dashed", color = "red")+ylim(c(0,50))
  return(list("mae_boxplot" = mae_boxplot, "mape_boxplot" = mape_boxplot))
}
```
Call boxplot_all function
```{r,boxplot call,echo=TRUE,warning=FALSE,message=FALSE}
bp<-boxplot_all()
bp$mae_boxplot
bp$mape_boxplot
```

###Confusion matrix
confusion_matrix function returns confusion matrix of polar and biometric in HR zone categorizing.
```{r setwd,echo=TRUE,warning=FALSE,message=FALSE}
confusion_matrix <- function(){
  ##create R dataframe from csv datafiles
  temp <- list.files(pattern="^PTek*")
  df <- data.frame()
  for (i in temp){
    ##substract subject id and activity type from filename
    tt<-str_split(i,'-')[[1]][3]
    Subject<-str_split(tt,' ')[[1]][2]
    Subject<-str_c("F-",Subject)
    Activity<-str_split(tt,' ')[[1]][3]
    ##label data with subject and activity
    t<-read.csv(i,header = T)
    t1<-t%>%mutate(Subject = Subject)%>%mutate(Activity = Activity)
    t1 <- subset(t1, select = -c(RRi,RRi.1))
    t1<-t1%>%slice(32:212)
    ##combine all files into one
    df <- bind_rows(t1,df)
  }
  ##calculate confusion matrix
  df$BiometRIC<-cut(df$HR.1,c(0,100,200),labels = c("Non-vigorous","Vigorous"))
  df$Polar<-cut(df$HR,c(0,100,200),labels = c("Non-vigorous","Vigorous"))
  df$BiometRIC <- factor(df$BiometRIC,levels=c("Vigorous","Non-vigorous"))
  df$Polar <- factor(df$Polar,levels=c("Vigorous","Non-vigorous"))
  cm<-confusionMatrix(df$BiometRIC,df$Polar,positive="Vigorous")
  return(cm)
}
```
Call confusion_matrix funcion, where reference is Polar H10, prediction is BiometRic
```{r cm, echo=TRUE,warning=FALSE,message=FALSE}
confusion_matrix()
```

###QoS
qos funtion returns two plots - mean absolute error vs qos, and data availability vs qos.
```{r qos,echo=TRUE,warning=FALSE,message=FALSE}
qos<-function(){
  ##create R dataframe from csv datafiles
  temp <- list.files(pattern="^PTek*")
  df <- data.frame()
  for (i in temp){
    ##substract subject id and activity type from filename
    tt<-str_split(i,'-')[[1]][3]
    Subject<-str_split(tt,' ')[[1]][2]
    Subject<-str_c("F-",Subject)
    Activity<-str_split(tt,' ')[[1]][3]
    ##label data with subject and activity
    t<-read.csv(i,header = T)
    t1<-t%>%mutate(Subject = Subject)%>%mutate(Activity = Activity)
    t1 <- subset(t1, select = -c(RRi,RRi.1))
    t1<-t1%>%slice(32:212)
    ##combine all files into one
    df <- bind_rows(t1,df)
  }
  
  ##category QoS data with 5% interval
  df$QoS_C <- cut(df$QoS, seq(0,100,5))
  df$QoS_F<-addNA(df$QoS_C)
  levels(df$QoS_F) <- c(0,levels(df$QoS_C))
  df$QoS_label <- factor(df$QoS_F, levels=levels(df$QoS_F),
    labels=c(seq(0,100,5)))
  df$QoS_num<-as.numeric(as.character(df$QoS_label))
  
  ##Making plot
  ###mape vs. QoS
  mae <- data.frame(Activity=c("sitting", "talking", "chewing", "drinking", "walking", "biking","running","housework"))
  for (n in c(0:19)){
    mae_cat<-df%>%filter(QoS_num >= 5*n)%>%group_by(Activity)%>%drop_na(HR.1,HR)%>%summarise(MAPE = MAPE(HR.1,HR)*100)
    mae<-left_join(mae,mae_cat,by="Activity")
    colnames(mae)[n+2]<-c(paste("",5*n,sep=""))
  }
  mae<-gather(mae,"QoS","MAPE",2:21)
  mae$MAPE<-as.numeric(mae$MAPE)
  mae$QoS<-as.factor(as.numeric(as.character(mae$QoS)))
  mae$Activity <- factor(mae$Activity,levels = c("sitting", "talking", "chewing", "drinking", "walking", "biking","running","housework"))
  mae_qos<-ggplot(mae,aes(QoS,MAPE,fill=Activity))+geom_bar(stat="identity")+facet_grid(Activity~.)+xlab("QoS level for retaining data")+ylab("MAPE (%)")+ggtitle("MAPE vs. QoS level for retaining data")+geom_hline(yintercept = 10,linetype="dashed", color = "red")
  
  ##DATA availability vs. QoS
  mae_count_before<-df%>%group_by(Activity)%>%drop_na(HR.1,HR)%>%summarise(count_before=length(HR.1))
  for (n in c(0:19)){
    mae_count_after<-df%>%filter(QoS_num >= 5*n)%>%group_by(Activity)%>%drop_na(HR.1,HR)%>%summarise(count_after=length(HR.1))
    mae_count_before<-left_join(mae_count_before,mae_count_after,by="Activity")
    mae_count_before[n+3]<- mae_count_before$count_after/mae_count_before$count_before*100
    colnames(mae_count_before)[n+3]<-c(paste("",5*n,sep=""))
  }
  
  mae_count<-gather(mae_count_before[-2],"QoS","DA",2:21)
  
  mae_count$DA<-as.numeric(mae_count$DA)
  mae_count$QoS<-as.factor(as.numeric(as.character(mae_count$QoS)))
  mae_count$Activity <- factor(mae_count$Activity,levels = c("sitting", "talking", "chewing", "drinking", "walking", "biking","running","housework"))
  
  da_qos<-ggplot(mae_count,aes(QoS,DA,fill=Activity))+geom_bar(stat="identity")+facet_grid(Activity~.)+xlab("QoS level for retaining data")+ylab("Data availability (%)")+ggtitle("Data Availability vs. QoS level for retaining data")+geom_hline(yintercept = 80,linetype="dashed", color = "red")
  return(list("mae"=mae_qos,"da"=da_qos))
}
```

Call qos function:
```{r qos call,echo=TRUE,warning=FALSE,message=FALSE}
qos<-qos()
qos$mae
qos$da
```
